Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-28 paper_jasa.tex
working on lwest pvalue picks

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{p_val_plot_signal.pdf}}
       \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$). The solid line is a 45 degree line which represents the equity shown in Equation \ref{eqn:power_estimate}. p-values are shown on square root scale scale. \blue{Is it better now?}}
    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
       \label{fig:pval_plot_signal}
\end{figure*}

@@

In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the observer would pick the plot in the lineup that had the lowest $p$-value. In this section we take a look at the picks that subjects made, to see if this is, indeed, true.

Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2} show the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively. In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizontally and the count of the number of subjects who selected this plot is displayed vertically. The dot represents this point, and the lines connect vertically with 0 count. Each row corresponds to a different $\beta$ value ordered top to bottom from lowest to highest. Columns correspond to the other levels of the experiment, sample size $(n)$, error standard deviation $(\sigma)$ and replicate. (Empty cells indicate no lineup shown with this combination.) Red indicates the actual data plot, which is not always the plot in the lineup with the lowest $p$-value. In both experiments as $\beta$ gets larger (stronger effect) people tended to select the plot with the lowest $p$-value. This is less so the case when $\beta$ is small, when standard deviation is large and sample size is small. The results are clearer for experiment 2, that used a continuous covariate. So for most observers, the assumption that they pick the plot with the smallest $p$-value would appear to be reasonable, and the actual power of the visual test should be close to the expected power.

@@

>>>>

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{p_val_plot_signal.pdf}}
       \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$). The solid line is a 45 degree line which represents the equity shown in Equation \ref{eqn:power_estimate}. p-values are shown on square root scale scale.}
    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
       \label{fig:pval_plot_signal}
\end{figure*}

@@

One assumption made in order to calculate the effect of lineup size in the calculations of visual $p$-value and signal strength was that subjects would tend to pick the plot in the lineup that had the strongest signal. In experiments 1 and 2, this corresponds to the plot with the smallest $p$-value. This can be examined for the data collected, to see if this assumption is, indeed, reasonable. %In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the observer would pick the plot in the lineup that had the lowest $p$-value. In this section we take a look at the picks that subjects made, to see if this is, indeed, true.

Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2} show the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively. In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizontally and the count of the number of subjects who selected this plot is displayed vertically. The dot represents this point, and the lines connect vertically with 0 count. Each row corresponds to a different $\beta$ value ordered top to bottom from lowest to highest. Columns correspond to the other levels of the experiment, sample size $(n)$, error standard deviation $(\sigma)$ and replicate. (Empty cells indicate no lineup shown with this combination.) Red indicates the plot with the lowest $p$-value in the lineup. In both experiments as $\beta$ gets larger (stronger effect) people tended to select the plot with the lowest $p$-value. This is less so the case when $\beta$ is small, when standard deviation is large and sample size is small. The results are clearer for experiment 2, that used a continuous covariate. But even when they didn't pick the plot with the lowest $p$-value they tended to oscillate their choices between the several low $p$-value plots. So for most observers, the assumption that they pick the plot with the smallest $p$-value would appear to be reasonable, and the actual power of the visual test should be close to the expected power.

There are some noticeable exceptions to this rule. In experiment 1, when $\beta=5, n=300, sd=12, rep=3$, people tended to pick the plot with the second smallest $p$-value.  Same experiment, when $\beta=0, n=100, sd=5, rep=1$ people overwhelming chose a plot with much larger $p$-value. For several of these exceptions, along with several easy lineups, we followed up with a new experiment using an eye-tracker to examine what the observers were looking at in making their choices. \cite{*** yifan's CC needs to be made into a tech report, and referenced here.}

@@

\green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice. Along the lines of the Army tests that Susan brought to the working group.}
